### 2019-03-01 Seneca.tex
mz - Michael's pass

\begin{figure}[t] \centering 
\includegraphics[scale=0.08]{Seneca}
\caption{The architecture of Seneca. 
User specifies hyperparameter options in the configuration file and provides lambda function. Seneca automatically package, deploy and optimize lambda function on AWS before completing grid search process and selecting best hyperparameter setting.
\label{fig:seneca}}
% \vspace{-0.2in}
\end{figure}

To facilitate model search and selection using the serverless architecture, 
we have developed Seneca, 
a framework for tuning the hyperparameters of machine learning applications 
in AWS Lambda. 
The Seneca pipeline consists of packaging, deployment, 
function optimization, and hyperparameter tuning. Figure~\ref{fig:seneca} 
shows the architecture of Seneca.  In the the upper-right-front, we
show the three inputs that Seneca expects from its users: a hyperparameter
configuration file, a dataset URL, and the path to 
the machine learning application.
The configuration file
specifies the range of values for each hyperparameter that the 
application expects.  Seneca uses this configuration to construct
the hyperparameter search space.  The dataset URL refers to a
valid dataset stored in the AWS Simple Storage 
Service (S3)\footnote{\url{https://aws.amazon.com/s3/}}.

Seneca automatically builds and deploys a AWS Lambda application
from the machine learning application specified.
To do so, Seneca launches a docker container that mirrors the
AWS Lambda execution environment, checks and installs
the machine learning application and any libraries it requires,
and compresses the application and uploads it to S3 (avoiding
the AWS Lambda handler size restriction of 10MB).
Seneca constructs an AWS Lambda handler from a template
that, when executed, will download and uncompress the application,
download the dataset and split it into a training and testing set,
and construct, test, and evaluate a model using the application
and a set of hyperparameter values passed in by Seneca as arguments.
Users can specify the train/test split that should be used by Seneca; the
default is 80\%/20\% for classification tasks.
The handler returns a testing score. 
Upon completion of this process, the container deploys the handler
to AWS Lambda using the
AWS Command line Interface\footnote{\url{https://aws.amazon.com/cli/}} 
and the developers credentials.

@@

The AWS Lambda pricing model is determined by the execution duration 
and memory of each invoked function. The cost of a function instance is referred to as 
the \texttt{compute charge}, which is the allocated memory 
multiplied by the
billed duration (execution time rounded up to the 
nearest 100ms)~\footnote{\url{https://aws.amazon.com/lambda/pricing/}}.
%https://aws.amazon.com/lambda/pricing/
To ensure that we keep the cost of Seneca as low as possible, we investigate
the impact of this pricing model on application execution and use
our findings to optimize cost.

Currently, allocated memory for a Lambda function can be set
from 128MB to 3008MB in increments of 64MB.
AWS documentation~\footnote{\url{https://docs.aws.amazon.com/lambda/latest/dg/resource-model.html}} 
states that Lambda 
allocates CPU in a way that is proportional to allocated memory size,
as is done for general purpose AWS EC2 instance types.

@@

To evaluate the impact of memory and CPU allocation on application cost,
we analyze a 3-D matrix multiple serverless benchmark~\cite{ref:matrix} using AWS Lambda.
We configure function instances to use the 46 possible allocated 
memory increments, invoke them, and record the billed duration and memory used by each.
Figure~\ref{fig:memory} shows the relationship between allocated memory 
(x-axis) and reciprocal of billed duration (y-axis). 
Figure~\ref{fig:compute_charge} shows the relationship between memory 
size (x-axis) and compute charge (y-axis). 
We observe that for this benchmark,
the CPU power plateaus after 1600 MB. 
Correspondingly, the compute charge declines from 128 MB to 1600 MB 
and then increases linearly.

These findings show that CPU does not increase proportionally with
an increase in allocated memory.  We use this relationship within
Seneca to optimize its cost (compute charge) via an extension
that enables it to automatically identify the appropriate
setting for allocated memory for each application.
However, instead of exhaustively testing all 46 possible 
memory configurations, which might be costly,  
Seneca employs the heuristic outlined in
Algorithm~\ref{algo:optimizer}.

The Seneca optimizer first configures and invokes the function using 
a user-defined payload.  From this run, Seneca obtains the memory 
used by function (reported by AWS Cloudwatch), 
which it uses as its starting point.
Seneca then defines two double-ended queues 
(\texttt{deque}) of length N, to store 
allocated memory and compute charge data of different
invocations. While the current 
allocated memory is less than or equal to 3008 MB, 
the optimizer reconfigures and invokes the function 
using the next increment for memory allocation.  It 
calculates the compute charge for each invocation 
using current allocated memory and billed duration.

We employ two exit conditions.  The first is when the 
compute charge monotonically increases across \texttt{deque}. 
The second is when the increase in slope is greater than a threshold. 
Once the optimizer finds that both conditions hold, 
it pops the left-most value from \texttt{deque} and configures 
the function to use that value for allocated memory for all future invocations.
After extensive experimentation, we find that 
N=5 and a slope threshold of 1 work well in most cases.
These values can be changed by users of the system however.
In addition, this optimization can be turned on or off via a command line
argument to Seneca.

\begin{algorithm}[]
\SetAlgoLined
\KwData{Typical payload}
\KwResult{Optimal allocated memory}
Find memory used by payload as starting point\;
Define deque for allocated memory \& compute charge\;
 \While{allocated memory $\leq$ 3008 MB}{

  \eIf{compute charge monotonically increases in deque \& slope $\geq$ 1}{
   Popleft from deque\;
   Configure allocated memory of function\; 
   Exit\;
   }{
   Increase allocated memory by 64 MB\;
   Probe lambda function\;
   Append allocated memory and compute charge to deque\;
  }
 }
 \caption{Seneca Optimizer Heuristic}
 \label{algo:optimizer}
\end{algorithm}

@@

For parallel function invocation, Seneca integrates 
Celery~\cite{ref:celery}. Celery is an asynchronous task queue 
that uses distributed message passing. Celery workers are processes 
that take tasks from the queue, execute the tasks with the arguments specified, 
and store the result that is returned 
in a database (we use Redis~\footnote{\url{https://redis.io/}} in our prototype).

Based on the configuration file, Seneca creates a list of 
payloads (function arguments) with
hyperparameter values.  
The complete list contains all combinations of the specified parameter
settings.  Seneca places the list on the queue and celery 
workers invoke the Lambda handler of the application, passing in the
payload (hyperparameter values) to the handler to use for model construction.  
Upon function
termination, the worker records the evaluation score (and serialized model)
in the database with the hyperparameter configuration from the payload.
When the queue is drained and all workers have completed, Seneca
extracts and reports the best score, configuration, and model from the database.
Users can then evaluate the prediction power of the models (using Seneca if desired) 
for other datasets (to amortize the time/cost of Seneca) without retraining.

We assume that the dataset supplied to Seneca by the user
is representative of datasets on which the 
resulting model will be used.  In addition, we use prediction error as the 
score (i.e., mean squared error for regression and accuracy percentage for classification) 
instead of $R^2$, which describes explanatory power, to avoid overfitting.
As part of future work, we are considering using multiple datasets and a ranges
of hyperparameter values (versus lists) to preclude the need for users to specify
them and to consider a wider range of values.

>>>>

\begin{figure}[t] \centering 
\includegraphics[scale=0.08]{Seneca}
\caption{The Seneca Architecture.
%Users specify hyperparameter options in the configuration file and provides lambda function. Seneca automatically package, deploy and optimize lambda function on AWS before completing grid search process and selecting best hyperparameter setting.
\label{fig:seneca}}
\vspace{-0.2in}
\end{figure}

To facilitate model search and selection using the serverless architecture, 
we have developed Seneca, 
a framework for tuning the hyperparameters of machine learning applications 
in AWS Lambda. 
The Seneca pipeline consists of packaging, deployment, 
function optimization, and hyperparameter tuning. Figure~\ref{fig:seneca} 
shows the architecture of Seneca.  In the the upper-right-front, we
show the three inputs that Seneca expects from its users: (\textit{A}) a hyperparameter
configuration file, (\textit{B}) a dataset URL, and (\textit{C}) the lambda function of machine learning application. The configuration file specifies the range of values for each hyperparameter that the application expects. Seneca applies Cartesian product to all options in this configuration to construct the search space. The dataset URL refers to a valid dataset stored in the AWS Simple Storage 
Service (S3)~\cite{ref:awss3}.
%\footnote{\url{https://aws.amazon.com/s3/}}.

Seneca automatically builds and deploys an AWS Lambda application from the machine learning application specified.
To do so, Seneca launches a docker container that mirrors the
AWS Lambda execution environment, checks and installs
the machine learning application and any libraries it requires,
and compresses the application and uploads it to S3 (a work-around
for the 10MB AWS Lambda function size restriction).
Seneca constructs an AWS Lambda function from a template
that, when executed, will download the dataset and split it into a training and testing set, and construct, test, and evaluate a model using the application
and a set of hyperparameter values passed in by Seneca as arguments.
Users can specify the train/test split ratio that should be used by Seneca; the
default is 80\%/20\% for classification tasks.
The function returns a testing score. 
Upon completion of this process, the container deploys the function
to AWS Lambda using the
AWS Command Line Interface~\cite{ref:awscli}
%\footnote{\url{https://aws.amazon.com/cli/}} 
and the developers credentials.

@@

The cost of using AWS Lambda (i.e. \texttt{compute charge}) 
is the billed duration (execution time rounded up to the
nearest 100ms)~\cite{ref:pricing}
%~\footnote{\url{https://aws.amazon.com/lambda/pricing/}}
multiplied by the allocated memory of each invoked function.
One goal of our work is to optimize memory use by these
applications in order to reduce cost, and to investigate
the trade offs of doing so.

Currently, allocated memory for a Lambda function can be set
from 128MB to 3008MB in increments of 64MB.
AWS documentation~\cite{ref:lambdalimits}
%~\footnote{\url{https://docs.aws.amazon.com/lambda/latest/dg/resource-model.html}} 
states that Lambda allocates CPU to functions proportionally to allocated memory size,
as is done for general purpose AWS EC2 instance types.

@@

To evaluate the relationship between memory, CPU, and cost, 
we analyze a 3-D matrix multiplication serverless benchmark~\cite{ref:matrix} using AWS Lambda.
We configure different functions to use each of the 46 possible allocated memory 
options.  Figure~\ref{fig:memory} shows the relationship between allocated memory 
(x-axis) and reciprocal of billed duration (y-axis). 
Figure~\ref{fig:compute_charge} shows the relationship between memory 
size (x-axis) and compute charge (y-axis). 
We observe that for this benchmark, billed duration 
plateaus after 1600 MB, at which point compute charge increases.
That is, we achieve no further execution time benefit (only cost increase)
after this point.

We use this relationship within
Seneca to optimize its cost (\texttt{compute charge}) via an extension
that enables it to automatically identify the appropriate
setting for allocated memory for each application.
However, instead of exhaustively testing all 46 possible 
memory configurations as we did for the matrix benchmark, 
which may be costly,  Seneca employs the heuristic outlined in
Algorithm~\ref{algo:optimizer}.

The Seneca optimizer first configures and invokes the function using a user-defined payload.  From this run, Seneca obtains the \texttt{maximum memory used}
by function reported by AWS Cloudwatch,
which it uses as its starting point.
Seneca then defines two double-ended queues 
(\textit{deque}) of length N, to store 
\texttt{allocated memory} and \texttt{compute charge} data of different invocations. While the current 
allocated memory is less than or equal to 3008 MB, 
the optimizer reconfigures and invokes the function 
using the next increment for memory allocation.  It 
calculates the compute charge for each invocation 
using current allocated memory and billed duration.

We employ two exit conditions. The first is when the  compute charge monotonically increases across \textit{deque}. The second is when the increase in slope is greater than a threshold. When the optimizer finds that both conditions hold, it pops the left-most value from \textit{deque} and configures the function to use that value for allocated memory for all future invocations. If these two conditions can not be satisfied through search, the allocated memory will be configured as the memory size that results in minimal compute charge within the \textit{deque}. After extensive experimentation, we find that $N=5$ and a slope threshold of 1 work best, but these values are configurable. In addition, this optimization can be turned on or off via a command line argument to Seneca.

\begin{algorithm}[]
\caption{Seneca Optimizer Heuristic}
\label{algo:optimizer}
\SetAlgoLined
\KwData{Typical payload}
\KwResult{Optimal allocated memory}
Find memory used by payload as starting point\;
Define deque for allocated memory \& compute charge\;
 \While{allocated memory $\leq$ 3008 MB}{
  \eIf{compute charge monotonically increases in deque \& slope $\geq$ threshold}{
   popleft from deque\;
   configure allocated memory as optimum\;
   exit()\;
   }{
   increase allocated memory by 64 MB\;
   probe lambda function\;
   append memory and compute charge to deque\;
  }
 }
\end{algorithm}

@@

To facilitate parallel function invocation, Seneca integrates 
Celery~\cite{ref:celery}.
%\footnote{\url{http://www.celeryproject.org/}}
Celery is an asynchronous task queue 
that uses distributed message passing. Celery workers are processes 
that take tasks from the queue, execute the tasks with the arguments specified, 
and store the result that is returned 
in a database (we use Redis~\cite{ref:redis}
%~\footnote{\url{https://redis.io/}} 
in our prototype).

Based on the configuration file, Seneca creates and enqueues a list of 
payloads (function arguments) for each combination of hyperparameter values.  
The Seneca celery workers invoke the application's Lambda function by each payload for model construction. Upon function termination, the worker records a score for the hyperparemeter configuration
in the database.  When the queue is drained and all workers have completed, Seneca extracts and reports the best score, configuration, and model from the database. Users can then use the model for inference given other datasets without retraining to amortize the time/cost of Seneca.
% Users can then evaluate the prediction power of the models (using Seneca if desired) 
% for other datasets without retraining to amortize the time/cost of Seneca.

We assume that the dataset supplied to Seneca by the user is representative of 
datasets on which the 
resulting model will be used.  In addition, we use prediction error as the 
score (i.e., mean squared error for regression and accuracy percentage for classification) 
instead of $R^2$, which describes explanatory power, to avoid overfitting.
As part of future work, we are considering using multiple datasets and a ranges
of hyperparameter values to preclude the need for users to specify
them and to consider a wider range of values.

### 2019-03-01 abstract.tex
mz - Michael's pass

The goal of our work is to simplify and expedite the 
construction and evaluation of machine learning models 
using autoscaled cloud computing systems and services.
To enable this, we develop an open source system, called Seneca, that leverages
the serverless programming model and its implementation in Amazon Web Services
(AWS) Lambda.  
Seneca takes a machine learning application, dataset, and a list of 
possible hyperparameter options.  It automatically constructs an AWS Lambda
handler that ingresses and splits the dataset into training
and testing subsets, and constructs, tests, and evaluates (i.e. scores) a machine
learning model for a given set of hyperparameter values.  Seneca concurrently invokes
functions for all combinations of the hyperparameters specified.
It then returns the configuration (or model) that results in the best score 
to the user.

We present the design and implementation of Seneca and introduce an extension to the
system that automatically optimizes memory use by the functions.  
Our empirical evaluation using multiple predictive machine learning applications 
for regression and classification shows that Seneca is able to quickly
identify the best performing model for the programs and datasets that we consider.
Moreover, its memory optimization reduces the cost of using Seneca 
by 10-30\% for the applications studied.

>>>>

The goal of our work is to simplify and expedite the 
construction and evaluation of machine learning models 
using autoscaled cloud computing systems and services.
To enable this, we develop an open source system, Seneca, that leverages
the serverless programming model and its implementation in Amazon Web Services
(AWS) Lambda.  
Seneca takes a machine learning application, dataset, and a list of 
possible hyperparameter options.  It automatically constructs an AWS Lambda
handler that ingresses and splits the dataset into training
and testing subsets, and constructs, tests, and evaluates (i.e. scores) a machine
learning model for a given set of hyperparameter values.  Seneca concurrently invokes
functions for all combinations of the hyperparameters specified.
It then returns the configuration (or model) that results in the best score 
to the user.

We present the design and implementation of Seneca and introduce an extension to the
system that automatically optimizes memory use by the functions.  
Our empirical evaluation using multiple predictive machine learning applications 
for regression and classification shows that Seneca is able to quickly
identify the best performing model for the programs and datasets that we consider.
Moreover, its memory optimization reduces the cost of using Seneca 
by 10--35\% for the applications studied.

### 2019-03-01 conclusion.tex
mz - Michael's pass

We present a new framework, called Seneca, for simplifying and expediting
the training and testing of machine learning models
using low cost cloud services.
Specifically, Seneca leverages the AWS Lambda for autoscaled
(elastic) and parallel execution of hyperparameter search 
and selection for machine learning models.  
Users provide Seneca with the application code and 
libraries, 1+ datasets, and the list 
of possible hyperparameter settings.  
Seneca uses this information to automatically
configure and deploy these functions in parallel for all possible 
combinations of hyperparameter values specified. Seneca
returns the best scoring model
and configuration to the user (for future use on other datasets).

We present the design, implementation, and cost optimization for Seneca.
The optimizer automatically optimizes function memory use to reduce the cost
of AWS Lambda use.  Our empirical evaluation using
multiple applications for regression and classification,
shows that Seneca is able to quickly
identify the best performing hyperparameter setting 
for the applications and datasets that we consider.  
We also show that Seneca memory optimization reduces the cost of using
Seneca by 10-30\% for the applications studied.

>>>>

We present a new framework, called Seneca, for simplifying and expediting
the training and testing of machine learning models in AWS Lambda.
%using low cost cloud services.
%Specifically, Seneca leverages the AWS Lambda for autoscaled
%(elastic) and parallel execution of hyperparameter search 
%and selection for machine learning models.  
Users provide Seneca with the application code and 
libraries, 1+ datasets, and the list 
of possible hyperparameter settings.  
Seneca uses this information to automatically
configure and deploy these functions concurrently for all possible 
combinations of hyperparameter values specified. Seneca
returns the best scoring model
and configuration to the user for future use on other datasets.

We present the design, implementation, and cost optimization for Seneca.
The optimizer automatically optimizes function memory use to reduce the cost
of AWS Lambda use.  Our empirical evaluation using
multiple applications for regression and classification,
shows that Seneca is able to quickly
identify the best performing hyperparameter setting 
for the applications and datasets that we consider.  
We also find that Seneca enables average speedups
of 294x for each additional dollar spent, and
that its memory optimization reduces the cost of using
Seneca by 10-35\% for the applications studied.

### 2019-03-01 intro.tex
mz - Michael's pass

To date however, clouds are used more for
enterprise services (object stores, databases, application servers, etc.)
than for elastic applications.
The reason for this is that it is challenging to configure complex
distributed systems for application use, 
and to leverage the auto-scaling that
clouds offer. To address this challenge, cloud providers 
have started to offer programming and execution environments
that preclude the need for server configuration,
under the \textit{serverless} moniker~\cite{ref:jonas2017occupy,ref:onesteptwostep,ref:peeking}.  
Serverless platforms automatically configure, manage, and scale applications
to significantly simplify cloud use.
%and broaden participation (by less expert
%users) in the use of cloud systems.
%xxx: http://shivaram.org/publications/pywren-socc17.pdf
%yyy: http://cidrdb.org/cidr2019/papers/p119-hellerstein-cidr19.pdf
%zzz: L. Wang, M. Li, Y. Zhang, T. Ristenpart, and M. Swift. “Peek- ing Behind the Curtains of Serverless Platforms”. In: USENIX ATC. Boston, MA: USENIX Association, 2018

Using the serverless model, application developers upload 
arbitrary computations in high level languages
as stateless functions to cloud-hosted, serverless platforms.
Functions are invoked (i.e. triggered)
automatically by the cloud in response to updates to other cloud services
(e.g. storage, queues, notification services, and API gateways, among others), which 
the developer configures. Serverless functions
communicate, persist, and access data only 
through their inputs or via shared, storage services.  
As a result, serverless applications are inherently elastic and can 
implement highly concurrent and parallel tasks.
In public clouds, users pay a small fee for the resources their 
functions use during execution, resulting in very low cost cloud use.
Although now available from all public cloud providers and as open 
source for private cloud systems, 
Amazon Web Services (AWS) Lambda\footnote{\url{https://docs.aws.amazon.com/lambda/latest/dg/lambda-dg.pdf}} %~\cite{ref:awslambdadg} 
was the first and is the most widely used
serverless public cloud platform.
%awslambda: https://docs.aws.amazon.com/lambda/latest/dg/lambda-dg.pdf

In this work, we investigate the efficacy of using AWS Lambda
for tuning machine learning applications.
To date, Lambda is not widely used for training and evaluating
machine learning models because of a concern that 
doing so will result in high overhead (i.e. be costly) because
of the stateless nature of serverless functions~\cite{ref:onesteptwostep}.
At the same time, identifying the ``best'' configuration for advanced
machine learning models is challenging given the large number of configuration
options (i.e. hyperparameters) typical for models today.
Hyperparameters govern the learning process of machine learning applications.
Given that parameter sweeps are embarrassingly parallel, we believe
that such tuning is a good fit for the serverless model.
To investigate this potential, its overhead, and to simplify the 
use of Lambda for training, testing, and evaluation of machine learning models, 
we design and develop a new system and toolset called \textit{Seneca}.

@@

We deploy Seneca on AWS Lambda and evaluate its tuning performance, cost,
and memory use for six machine learning applications and datasets. We 
find that Seneca is fast, inexpensive, and effective for model 
construction and comparison. Seneca is also able to identify
automatically the
best memory configuration for each application, further lowering its cost
by 10-30\%.
We intend to make Seneca, its applications, and their datasets publicly available
when/if this paper is accepted for publication.  In the sections
that follow, we overview our design and implementation of Seneca and then
present our empirical methodology and results.

\ignore{
Recent work~\cite{ref:onesteptwostep} argues however that AWS Lambda

Serverless~\cite{ref:serverless} computing is a rising function-based (also known as Functions-as-a-Service~\cite{ref:faas}) programming and deployment paradigm on top of cloud infrastructure, in which programmers manifest business logic without concerning provisioning servers by modular functions that are triggered by incoming events and return computation results in serialized format like XML or JSON. Such events include HTTP requests from clients, possibly proxied by API Gateway, or communications between cloud services based on heterogeneous protocols. Depending on the supported runtimes from cloud vendors, functions could be written in various high-level languages (i.e. NodeJS, Python, Java, etc.) and take advantage of serverless SDK to interact with other cloud services. Until recently, the leading serverless platform, Amazon Web Services(AWS) Lambda, expands language support by custom runtime~\cite{ref:customruntime} where software engineers can implement functions by any programming languages.

Primarily, there are four evolutionary advantages that serverless architecture brings to cloud computing. First and foremost, it raises the abstraction layer from virtual machines to function containers~\cite{ref:container} that abstracts away the responsibilities of infrastructure management from DevOps~\cite{ref:devops} personnel to the cloud provider, in terms of scaling, patching, provisioning, error-handling and terminating underlying bare metal servers or clusters. Such shift allows more agile development life cycle and more product-oriented programming iteration. Secondly, the event-driven and paralleled container runtime of serverless function provides more fine-grained computational resource isolation and usage. Hence, serverless architecture makes cloud operation more secure and energy-efficient compared to traditional virtual machine instance runtime. Thirdly, serverless architecture decomposes monolithic application code base even further from microservices~\cite{ref:microservices} into granular functions, which are easier to be created, maintained and refactored by team or individual programmer. Furthermore, Function-as-a-Service applications could be autoscaled independently with respect to modular functions depending on the volume of request. This capability also contributes to the energy efficiency and eliminating throttling bottlenecks. Lastly, serverless computing provides "Pay-per-execution" pricing model~\cite{ref:pricing}. In a real world case study~\cite{ref:serverless_econ}, the server operational cost on serverless platform could be reduced dramatically, given a relatively lower amount of Transactions Per Second (TPS). According to these merits, the stateless, event-driven and embarrassingly parallel tasks are potentially cost-effective with lower latency by running on serverless framework. 
}

>>>>

To date however, clouds are used more for
enterprise services (object stores, databases, application servers, etc.)
than for elastic applications.
The reason is that it is challenging to configure complex
distributed systems for application use, 
and to leverage the auto-scaling that
clouds offer. To address this challenge, cloud providers 
have started to offer programming and execution environments
that preclude the need for server configuration,
under the \textit{serverless} moniker~\cite{ref:jonas2017occupy,ref:onesteptwostep,ref:peeking}.  
Serverless platforms automatically configure, manage, and scale applications
to significantly simplify cloud use.

Using the serverless model, application developers upload arbitrary computations in high level languages as stateless functions to cloud-hosted, serverless platforms, where functions are triggered automatically by the cloud in response to updates from other cloud services (e.g. storage, queues, notification services, and API gateways, 
among others). Serverless functions must execute under a time bound (e.g. 15 minutes) and an allocated memory size 
(e.g. 3GB) or else the platform will terminate the function. They
communicate, persist, and access data only 
through their inputs or via shared storage services. As a result, serverless applications are inherently elastic and can 
implement highly concurrent and parallel tasks.
In public clouds, users pay a small fee for the resources their 
functions use during execution, resulting in very low cost cloud use.
Although now available from all public cloud providers and as open 
source for private cloud systems, 
Amazon Web Services (AWS) Lambda~\cite{ref:awslambdadg}
%\footnote{\url{https://docs.aws.amazon.com/lambda/latest/dg/lambda-dg.pdf}}
was the first and is the most widely used
serverless public cloud platform.
%awslambda: https://docs.aws.amazon.com/lambda/latest/dg/lambda-dg.pdf

In this work, we investigate the efficacy of using AWS Lambda
for tuning machine learning applications in parallel.
To date, Lambda is not widely used for training and evaluating
machine learning models because of a concern that 
doing so will result in high overhead (i.e. be costly) because
of the stateless nature of serverless functions~\cite{ref:onesteptwostep}.
At the same time, identifying the ``best'' configuration for advanced
machine learning models is challenging given the large number of configuration
options (i.e. hyperparameters) typical for models today.
Hyperparameters govern the learning process of machine learning applications.
Given that parameter sweeps are embarrassingly parallel, we believe
that such tuning is a good fit for the serverless model.
To investigate this potential, its overhead, and to simplify the 
use of Lambda for training, testing, and evaluation of machine learning models, 
we design and develop a new system and toolset called \textit{Seneca}.

@@

We deploy Seneca on AWS Lambda and evaluate its tuning performance, cost,
and memory use for five machine learning applications and datasets. We 
find that Seneca is fast, inexpensive, and effective for model 
construction and comparison. Seneca is also able to identify
automatically the
best memory configuration for each application, further lowering its cost
by 10-35\%. Relative to execution in AWS EC2, Seneca enables average speedups of 294x for each additional dollar 
spent for the applications and datasets we consider.
We intend to make Seneca, its applications, and their datasets publicly available
when/if this paper is accepted for publication.  
We next overview our design and implementation of Seneca and then
present our empirical methodology and results.

### 2019-03-01 related_work.tex
mz - Michael's pass

We next overview related work. We discuss contributions that focus
on serverless computing, automatic packaging, 
and machine learning (ML) model optimization.

Multiple researchers have investigated the efficacy and overhead of the
serverless programming model and 
implementations~\cite{ref:jonas2017occupy,ref:onesteptwostep,ref:baldini2017,ref:lin2018tracking}.
%The authors in \cite{ref:onesteptwostep} evaluate the cost and performance of using
%serverless for data-intensive computing.  
They identify some weaknesses in AWS
Lambda for data-intensive computing, including the inability to 
effectively train models.
Our work shows that it is possible and cost effective to use
AWS Lambda to train models for the subset of ML applications that we 
consider.

PyWren~\cite{ref:jonas2017occupy} shows that stateless functions
are a natural fit for distributed computing models. The technique
abstracts away cluster management overhead and is ideal for 
embarrassingly parallel jobs.  ExCamera~\cite{ref:encoding} presents a 
framework for running general-purpose parallel tasks (encoding 4K video) 
on a commercial serverless platform using multithreading. 
Cirrus~\cite{ref:jonas2019cloud} attempts to train ML models 
using a parameter server and serverless functions. 
An empirical comparison of resource utilization and performance isolation
is provided in~\cite{ref:peeking} for AWS Lambda, 
Azure Functions, and Google Cloud Functions.

The serverless framework~\cite{ref:serverless_framework} provides
automated packaging and deployment for serverless functions across clouds. 
The serverless framework uses CloudFormation\footnote{\url{https://aws.amazon.com/cloudformation/}} for deployment in AWS Lambda, which introduces additional cost. 
The cloud infrastructure provisioning 
framework Terraform\footnote{\url{https://www.terraform.io/}}
also provides automated deployment of functions to serverless platforms
(however, without replicating the Lambda runtim, introducing deployment errors).
Seneca uses a local docker container to avoid this cost and overhead. 
%In addition, AWS CodeDeploy~\cite{ref:codedeploy} serves as component in the continuous integration and delivery pipeline for serverless architectures, but it also introduces monetary cost overhead to do so.

Automated hyperparameter optimization is the focus of
many projects. Google Vizier~\cite{ref:vizier} provides a service for 
black-box optimization. Optunity~\cite{ref:claesen2014hyperparameter} 
and Hyperopt\footnote{\url{http://hyperopt.github.io/hyperopt/}}
provide a Python library for hyperparameter tuning. 
%This can be considered as precursory work of Hyperopt~\cite{ref:hyperopt}, 
%which is a more advanced Python library for serial and parallel hyperparameter 
%optimization over awkward search space. 
Hyperas\footnote{\url{http://maxpumperla.com/hyperas/}} adds another 
abstraction layer to hyperopt to facilitate hyperparameter tuning for 
Keras~\footnote{\url{https://keras.io/}}. However, no extant work of which we are aware leverages
serverless to perform hyperparameter tuning and memory optimization in parallel 
for ML applications.

>>>>

As related work, we consider recent advances
in evaluating serverless computing for different application domains, 
automatic deployment for serverless, and machine learning (ML) model optimization. 
For the former, much work has investigated the efficacy and overhead of the
serverless programming model and 
implementations~\cite{ref:jonas2017occupy,ref:onesteptwostep,ref:baldini2017,ref:lin2018tracking}.
%The authors in \cite{ref:onesteptwostep} evaluate the cost and performance of using
%serverless for data-intensive computing.  
The authors identify challenges with 
using AWS Lambda to train machine learning (ML) models.
Our work however, shows that it is possible to leverage the concurrency 
and parallelism in AWS Lambda to perform fast grid search for the
subset of ML applications that we consider.

PyWren~\cite{ref:jonas2017occupy} uses serverless for
different distributed computing models. The technique abstracts away cluster management overhead and is ideal for embarrassingly parallel jobs.
ExCamera~\cite{ref:encoding} presents a framework for running general-purpose parallel tasks (encoding 4K video) on a commercial serverless platform using multithreading. Cirrus~\cite{ref:jonas2019cloud} attempts to train ML models using a parameter server and serverless functions. 
%An empirical comparison of resource utilization and performance isolation
%is provided in~\cite{ref:peeking} for AWS Lambda, 
%Azure Functions, and Google Cloud Functions.

The serverless framework~\cite{ref:serverless_framework} provides
automated packaging and deployment for serverless functions across clouds. 
GammaRay~\cite{gammaray18} does so for AWS Lambda to insert profiling instrumentation.
The serverless framework uses 
CloudFormation~\cite{ref:cloudformation} for deployment in AWS Lambda, which introduces additional cost. 
The cloud infrastructure provisioning 
framework Terraform~\cite{ref:terraform}
also provides automated deployment of functions to serverless platforms.
%(however, without replicating the Lambda runtime, introducing deployment errors).
Seneca uses a local docker container to avoid  cost and overhead (vs these related
works), which guarantees execution compatibility for AWS Lambda. 
%In addition, AWS CodeDeploy~\cite{ref:codedeploy} serves as component in the continuous integration and delivery pipeline for serverless architectures, but it also introduces monetary cost overhead to do so.

Automated hyperparameter tuning is the focus of
many projects. Google Vizier~\cite{ref:vizier} provides a service for 
black-box optimization. Optunity~\cite{ref:claesen2014hyperparameter} 
and Hyperopt~\cite{ref:hyperopt}
provide a Python library for hyperparameter tuning. 
%This can be considered as precursory work of Hyperopt~\cite{ref:hyperopt}, 
%which is a more advanced Python library for serial and parallel hyperparameter 
%optimization over awkward search space. 
Hyperas~\cite{ref:hyperas} adds another 
abstraction layer to hyperopt to facilitate hyperparameter tuning for 
Keras~\cite{ref:keras}. However, no extant work of which 
we are aware leverages
serverless to perform hyperparameter tuning and memory optimization in parallel 
for ML applications.
